\input{../../template.ltx}
\usepackage{amsmath}
\usepackage{multicol}
\begin{document}

\osuetitle{2}

\section*{Assignment -- Fork Fourier Transform}
Implement the Cooley-Tukey Fast Fourier Transform\footnote{\url{https://en.wikipedia.org/wiki/Cooley-Tukey\_FFT\_algorithm}} algorithm.
\begin{verbatim}
    SYNOPSIS
        forkFFT
\end{verbatim}

\subsection*{Instructions}
The input of the program are floating point values, which should be read from \osueglvar{stdin}.
Subsequent values are separated by a newline character.
The sequence ends when an EOF (End Of File) is encountered.

Your program must accept any number of values.
Terminate the program with exit status \verb|EXIT_ERROR|
if an invalid input is encountered.

The program computes the Fourier Transform of its input values recursively,
i.e. by calling itself:
\begin{enumerate}
\item If the array consists of only 1 number, write that number to \osueglvar{stdout} and exit.
\item Otherwise the array consists of $n>1$ numbers.
Split it into two parts with $n/2$ numbers each,
such that one part contains all the numbers at even indices
and the second part contains all the numbers at odd indices:
\[
P_E=\{A[0],\;A[2],\;A[4],\;\dots\}
\]
\[
P_O=\{A[1],\;A[3],\;A[5],\;\dots\}
\]
Terminate the program with exit status \verb|EXIT_ERROR|
if the length of the array is not even.

\item Using \osuefunc{fork(2)} and \osuefunc{execlp(3)},
recursively execute this program in two child processes,
one for each of the two parts.
Use two unnamed pipes per child
to redirect \osueglvar{stdin} and \osueglvar{stdout}
(see \osuefunc{pipe(2)} and \osuefunc{dup2(3)}).
Write $P_E$ to \osueglvar{stdin} of one child
and $P_O$ to \osueglvar{stdin} of the other child.
Read the respective results from each child's \osueglvar{stdout}.
The two child processes must run simultaneously!

\item Use \osuefunc{wait(2)} or \osuefunc{waitpid(2)}
to read the exit status of the children.
Terminate the program with exit status \verb|EXIT_ERROR|
if the exit status of any of the two child processes is not \verb|EXIT_SUCCESS|.

\item Let $R_E$ be the result of the even part $P_E$
and $R_O$ be the result of the odd part $P_O$.
The result $R$ of the Fourier Transform of the entire array can now be computed
by applying the ``butterfly'' operation to the two results:
\[
\forall k\;|\;0\leq k<n/2\;:\quad R[k]=R_E[k]+e^{-\frac{2\pi i}{n}\cdot k}\cdot R_O[k]
\quad\text{and}\quad R[k+n/2]=R_E[k]-e^{-\frac{2\pi i}{n}\cdot k}\cdot R_O[k]
\]
which is identical to:
\[
\forall k\;|\;0\leq k<n/2\;:\quad
\renewcommand{\arraystretch}{1.8}
\begin{array}{r@{\;=\;}l}
 R[k]     & R_E[k]+\left(\text{cos}(-\frac{2\pi}{n}\cdot k)+i\cdot\text{sin}(-\frac{2\pi}{n}\cdot k)\right)\cdot R_O[k]\\
R[k+n/2] & R_E[k]-\left(\text{cos}(-\frac{2\pi}{n}\cdot k)+i\cdot\text{sin}(-\frac{2\pi}{n}\cdot k)\right)\cdot R_O[k]
\end{array}
\]
You may approximate the value of $\pi$ with $3.141592654$.
\item The resulting array is printed to \osueglvar{stdout}, one value per line.
Since the resulting numbers are complex, the program prints two values per line:
the real and the imaginary part of each value, separated by a whitespace
(likewise $R_E$ and $R_O$ are imaginary,
thus when reading them from the child processes,
two values per line must be parsed).
You may append \verb|*i| or similar to the second value
to indicate that it is the imaginary part (see examples).

\end{enumerate}

Throughout the program,
it is sufficient to use single precision floating point numbers.

\subsection*{Hints}
\begin{enumerate}
\item Use \osuefunc{strtof(3)} to parse floating point values.
The \verb|endptr| argument can be helpful for reading two numbers per line
when reading from the child processes.
\item Use the header \osueglvar{math.h} for the calculation of the sine and cosine.
Add \verb|-lm| to the linker options to use these functions.
\item When dealing with complex numbers,
you may find following property useful to implement multiplication:
$(a+i\cdot b)(c+i\cdot d)=a\cdot c-b\cdot d+i\cdot(a\cdot d+b\cdot c)$
Addition should be straight-forward to implement
and no other operation with complex numbers is required.
\item In order to avoid endless recursion\footnote{\url{http://en.wikipedia.org/wiki/Fork\_bomb}},
fork only if the input number is greater than 1.
\item To output error messages and debug messages, always use
\osueglvar{stderr} because \osueglvar{stdout} is redirected in most cases.
\end{enumerate}

\subsection*{Examples}
\begin{multicols}{2}
\paragraph{Constant signal:}
\begin{verbatim}
$ cat constant.txt
1.0
1.0
$ ./forkFFT < constant.txt
2.0 0.0*i
0.0 0.0*i












\end{verbatim}

\paragraph{Sinus wave:}
\begin{verbatim}
$ cat sine.txt
0.000000
0.707107
1.000000
0.707107
0.000000
-0.707107
-1.000000
-0.707107
$ ./forkFFT < sine.txt
0.000000 0.000000*i
0.000000 -4.000001*i
0.000000 0.000000*i
0.000000 -0.000001*i
0.000000 0.000000*i
-0.000000 0.000001*i
0.000000 0.000000*i
-0.000000 4.000001*i
\end{verbatim}
\end{multicols}

\osueguidelinestwo

\end{document}
